Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2019-02-11 hpmoon-asoco.tex
Addressing R2 changes for issues #28 #29 #31

\begin{bluecolor}
One of the  most extensively used methods to deal with high-dimensional, multi-objective optimization problems are distributed co-evolutionary algorithms. 
However, only
decomposable problems could strictly be implemented in parallel this
way. 
A new parallel island-model evolutionary algorithm approach, which we call {\em selective operator application}, is used here to deal with un-decomposable
problems: 
instead of sharing parts
of individuals, each island, hosted by a single processor, applies the variation operators to a
specific subset of the whole individual. This avoids extra steps to recompose the complete individuals from other islands before evaluation.

 In this paper, we intend to prove that automatically setting the size of these
 overlapping sections, since it focuses the search between borders of the different search spaces, is able to obtain better
 results than using the same size independently of the number of islands. In order to prove that, we
 have compared this method to other parallel evolutionary
techniques considering different number of islands, chromosome sizes
and benchmarks. The analysis of the obtained experimental results, by
using different metrics, shows that our approach can provide
statistically significant improvements with respect to the base
algorithm in high-dimensional, un-decomposable, multi-objective problems.

Results also show that the relation between the number of
islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables: the higher
the dimensionality and the overlapping, the greater the benefit of distribution in a
large number of islands.
This opens up a new challenge to be addressed in the future, by
experimenting with new ways to calculate the overlapping sizes depending
of different factors and parameters.
\end{bluecolor}

@@

%1. MOP ARE
Multi-objective optimization problems (MOP) are those where several
objectives have to be simultaneously optimized
\citep{Mora13paretobased,LIU2017344}. Solving a MOP implies optimizing a function composed of several independent cost functions, one per objective. In
these problems the aim is to obtain a set of solutions that are better
than the rest, considering all the objectives; this set is known as
the Pareto Front (PF). The solutions in this set are {\em non-dominated},
which means that there is no other solution that is equal or better
for all objectives; the objective of MOP algorithms is to find as many
solutions equally distributed in the Pareto Front as possible, and in order to do that the
exploration of the search space must be very efficient, trying to
cover many possible exploration paths at the same time.

@@

%The availability of metaheuristics which are able to efficiently approach these multi-objective optimization problems have made it possible to cope with many real applications from such multi-objective point of view [4, 5, 6, 7, 8].  Evolutionary algorithms (EAs) [9] are bio-inspired meta-heuristics that can be effectively used to find nearly optimal solutions for optimization problems and have been extended to multi-objective optimization through quite a number of Multi-Objective Evolutionary Algorithms (MOEAs) [29].

@@

\textcolor{blue}{EAs have been extended to multi-objective optimization through quite a number of Multi-Objective Evolutionary Algorithms (MOEAs) \cite{TalbiUnified2018}}.
One of the best known and referenced MOEAs
\citep{Dorronsoro13superlinear} is the {\em 
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, which are divided into different ranks, based on the
dominance, to be selected for the next generations.
% This paragraph has nothing to do with anything. Is it useful for
% what you are doing in the paper? Is there any challenge that you
% will be solving here? For instance, you could talk about the problem
% it's got for parallelization and how it deals with them - JJ FERGU: Could you add some paragraph to justify it? (use \textcolor{blue}{} please)

@@

\textcolor{blue}{Real-world MOPs usually require a high number of decision variables \cite{Zhang17DECAL,XU2018268}}, 
meaning that the
MOEAs need to deal with large individuals and spend significant extra
time for crossover, mutation and migration. %depending on the
                                %language, this time might not be
                                %significant at all, and will depend
                                %on the length of the chromosome, not
                                %the part of the chromosome you will
                                %be applying it to. Besides, this is
                                %not the critical part of the
                                %algorithm, the critical part is
                                %searching the solution space. 
Different authors have % this is only one possibility. You should talk
                       % about simple island models, too - JJ
proposed methods to divide the decision space (the chromosome) to
improve the performance and solution quality. In this aspect, the
co-evolution model is a dimension-distributed model where a % first
                                % reference to co-evolution in this
                                % paper. Is it still a paper about
                                % co-evolution? - JJ
high-dimensional problem is divided into lower dimensional ones
\citep{Gong15models,Tonda12cooperative}, and evolved separately. One
example of application of this technique was described in
\citep{Kimovski15Parallel}. The method presented in that work involves
different workers that evolve sub-populations created and recombined
by a master process, which performs different recombination
alternatives from the parts returned by worker processes. A high
dimensional problem was used to compare these 
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

A similar approach used to solve this kind of problems is the {\em
  SeLective operator Application} (SLA) \textcolor{blue}{presented in this paper}. In this case, each island
deals with the whole chromosome, but only modifies a fragment of it in
the crossover and mutation phase depending on the number of islands,
using the whole chromosome for fitness calculation. This allows to
deal with un-decomposable problems. % if the whole chromosome is used,
                                % mutation and crossover will take the
                                % same time. You are modifying the
                                % search space, though. 
In our previous work \citep{Garcia16hpmoon} we used this method in a
preliminary experimental setup, although it is worth mentioning we are
establishing its name and acronym in this paper for the first time.
We proved 
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. In that paper, we showed that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. %where we couldn't prove that... Or, we
                          %couldn't experiment... - JJ
This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically \textcolor{blue}{sets} the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions.

% We have to
                                % emphasize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ FERGU: rewritten previous paragraph

@@

>>>>

\begin{bluecolor}

%One of the  most extensively used methods to deal with high-dimensional, multi-objective optimization problems are distributed co-evolutionary algorithms. However, only decomposable problems could strictly be implemented in parallel this way. A new parallel island-model evolutionary algorithm approach, which we call {\em selective operator application}, is used here to deal with un-decomposable problems: instead of sharing parts of individuals, each island, hosted by a single processor, applies the variation operators to a specific subset of the whole individual. This avoids extra steps to recompose the complete individuals from other islands before evaluation.

%In this paper, we intend to demonstrate that automatically setting the size of these  overlapping sections, since it focuses the search between borders of the different search spaces, is able to obtain better  results than using the same size independently of the number of islands. In order to demonstrate that, we  have compared this method to other parallel evolutionary techniques considering different number of islands, chromosome sizes and benchmarks. The analysis of the obtained experimental results, by using different metrics, shows that our approach can provide statistically significant improvements with respect to the base algorithm in high-dimensional, un-decomposable, multi-objective problems.

%Results also show that the relation between the number of islands (subpopulations) and the length of the chromosome (number of decision variables) is also a relevant factor to determine the most efficient alternative to distribute the decision variables: the higher the dimensionality and the overlapping, the greater the benefit of distribution in a large number of islands. This opens up a new challenge to be addressed in the future, by experimenting with new ways to calculate the overlapping sizes depending of different factors and parameters.

%%%NEW AND SHORTED
Parallel island-model co-evolutionary algorithms are a well-known and suitable methods to deal with large multi-objective optimization problems. In these algorithms each island modifies a fragment of the chromosome that encodes a possible solution to the problem. In this paper, we want to demonstrate that automatically setting the size of the overlapping fragments depending on the number of islands obtains better results than using the same size. We  have compared this method to other parallel evolutionary techniques considering different number of islands, chromosome sizes and benchmarks. The analysis of the obtained experimental results, by using different metrics, shows that our approach can provide statistically significant improvements with respect to the base algorithm in high-dimensional, un-decomposable, multi-objective problems. This opens a very challenging line to automatically adapt the overlapping sizes.

\end{bluecolor}

@@

\begin{bluecolor}
One of the  most extensively used methods to deal with high-dimensional, Multi-Objective Optimization problems (MOPs) are distributed co-evolutionary algorithms \cite{Gong15models}. However, only decomposable problems could strictly be implemented in parallel this way. A new parallel island-model evolutionary algorithm approach \cite{Alba02Parallelism}, is used here to deal with un-decomposable problems: instead of sharing parts of individuals, each island, hosted by a single processor, applies the variation operators to a specific subset of the whole individual. This avoids extra steps to recompose the complete individuals from other islands before evaluation.
\end{bluecolor}

%1. MOP ARE
MOPs are those where several
objectives have to be simultaneously optimized
\citep{Mora13paretobased,LIU2017344}. Solving a MOP implies optimizing a function composed of several independent cost functions, one per objective. In
these problems the aim is to obtain a set of solutions that are better
than the rest, considering all the objectives; this set is known as
the Pareto Front (PF). The solutions in this set are {\em non-dominated},
which means that there is no other solution that is equal or better
for all objectives; the objective of MOP algorithms is to find as many
solutions equally distributed in the Pareto Front as possible, and in order to do that the
exploration of the search space must be very efficient, trying to
cover many possible exploration paths at the same time.

@@

@@

EAs have been extended to multi-objective optimization through quite a number of Multi-Objective Evolutionary Algorithms (MOEAs) \cite{TalbiUnified2018}.
One of the best known and referenced MOEAs
\citep{Dorronsoro13superlinear} is the {\em 
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, which are divided into different ranks, based on the
dominance, to be selected for the next generations.

@@

\textcolor{blue}{Real-world MOPs usually require a high number of decision variables \cite{Zhang17DECAL,XU2018268}}, 
meaning that the
MOEAs need to deal with large individuals and spend significant extra
time for crossover, mutation and migration.
Different authors have 
proposed methods to divide the decision space (the chromosome) to
improve the performance and solution quality. In this aspect, the
co-evolution model is a dimension-distributed model where a 
high-dimensional problem is divided into lower dimensional ones
\citep{Gong15models,Tonda12cooperative}, and evolved separately. One
example of application of this technique was described in
\citep{Kimovski15Parallel}. The method presented in that work involves
different workers that evolve sub-populations created and recombined
by a master process, which performs different recombination
alternatives from the parts returned by worker processes. A high
dimensional problem was used to compare these 
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

A similar approach used to solve this kind of problems is the {\em
  SeLective operator Application} (SLA) \textcolor{blue}{presented in this paper}. In this case, each island
deals with the whole chromosome, but only modifies a fragment of it in
the crossover and mutation phase depending on the number of islands,
using the whole chromosome for fitness calculation. This allows to
deal with un-decomposable problems.

In our previous work \citep{Garcia16hpmoon} we used this method in a
preliminary experimental setup, although it is worth mentioning we are
establishing its name and acronym in this paper for the first time.
We proved 
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. In that paper, we showed that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. 
This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically \textcolor{blue}{sets} the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions.

@@

\textcolor{blue}{Results also show that the relation between the number of islands (subpopulations) and the length of the chromosome (number of decision variables) is also a relevant factor to determine the most efficient alternative to distribute the decision variables: the higher the dimensionality and the overlapping, the greater the benefit of distribution in a large number of islands. This opens up a new challenge to be addressed in the future, by experimenting with new ways to calculate the overlapping sizes depending of different factors and parameters.}
